Quantum Hoare Type Theory

### 2020-06-08 ms-paper/ms-paper.tex
Remove conditional and move to future work
WIP

\begin{figure}[ht]
	\begin{tabular}{lrcl}
		\textit{Types} & $A, B, C$ & ::= &
		\begin{minipage}[t]{0.45\columnwidth}%
			$ \type{1} \mid \type{Bool} \mid \type{Qbit} \mid \type{U} \mid \type{Pure} \mid A \otimes B \mid \Pi x{ : }A.B \mid \Delta. \Psi .\{P\} x{ : }A \{Q\}$
		\end{minipage}\\ \\
		\textit{Assertions} & $P, Q, R$ & ::= &
		\begin{minipage}[t]{0.45\columnwidth}%
			$ \top \mid \bot \mid P \wedge Q \mid P \vee Q \mid P \supset Q \mid \neg P \mid \exists x{:}A.P \mid \forall x{:}A.P \mid \exists h{:}\kw{heap}.P \mid \forall h{:}\kw{heap}.P \mid \kw{Id}_A(M, N) \mid \kw{HId}(H, G) \mid \kw{indom}(H,M) $
		\end{minipage}\\ \\
		\textit{Q. Heaps} & $H, G$ & ::= & $ h \mid \kw{empty} \mid \kw{upd}(H, M, N) $\\ \\
		\textit{Elim terms} & $K, L$ & ::= & $ x \mid K M \mid M : A $\\
		\textit{Intro terms} & $M, N, O$ & ::= & $ K \mid () \mid \lambda x.M \mid \kw{do} E \mid \kw{true} \mid \kw{false} $\\ \\
		\textit{Q. Commands} & $c$ & ::= &
		\begin{minipage}[t]{0.45\columnwidth}%
			$ \kw{mkQbit} M \mid \kw{measQbit} M \mid \kw{applyU} M \mid \kw{if} M \kw{then} N \kw{else} O$
		\end{minipage}\\
		\textit{Computations} & $E, F$ & ::= & $ \kw{return} M \mid x \leftarrow K; E \mid x \Leftarrow c; E \mid x =_A M; E $\\ \\
		\textit{Variable context} & $\Delta$ & ::= & $ \cdot \mid x : A, \Delta $ \\
		\textit{Q. Heap context} & $\Psi$ & ::= & $ \cdot \mid h, \Psi$ \\
		\textit{Assertion context} & $\Gamma$ & ::= & $ \cdot \mid P, \Gamma$ \\
	\end{tabular}
	\Description{Syntax of Quantum Hoare Type Theory}
	\caption{Syntax of Quantum Hoare Type Theory}
	\label{fig:syntax}
\end{figure}

@@

\paragraph{Quantum Commands}
These are the quantum-specific effectful commands from QIO for initialization, unitary application and measurement of qubits. We also support classical control using the \texttt{if} $M$ \texttt{then} $N$ \texttt{else} $O$ construct similar to HTT.

@@

@@

@@

>>>>

\begin{figure}[ht]
	\begin{tabular}{lrcl}
		\textit{Types} & $A, B, C$ & ::= &
		\begin{minipage}[t]{0.45\columnwidth}%
			$ \type{1} \mid \type{Bool} \mid \type{Qbit} \mid \type{U} \mid \type{Pure} \mid A \otimes B \mid \Pi x{ : }A.B \mid \Delta. \Psi .\{P\} x{ : }A \{Q\}$
		\end{minipage}\\ \\
		\textit{Assertions} & $P, Q, R$ & ::= &
		\begin{minipage}[t]{0.45\columnwidth}%
			$ \top \mid \bot \mid P \wedge Q \mid P \vee Q \mid P \supset Q \mid \neg P \mid \exists x{:}A.P \mid \forall x{:}A.P \mid \exists h{:}\kw{heap}.P \mid \forall h{:}\kw{heap}.P \mid \kw{Id}_A(M, N) \mid \kw{HId}(H, G) \mid \kw{indom}(H,M) $
		\end{minipage}\\ \\
		\textit{Q. Heaps} & $H, G$ & ::= & $ h \mid \kw{empty} \mid \kw{upd}(H, M, N) $\\ \\
		\textit{Elim terms} & $K, L$ & ::= & $ x \mid K M \mid M : A $\\
		\textit{Intro terms} & $M, N, O$ & ::= & $ K \mid () \mid \lambda x.M \mid \kw{do} E \mid \kw{true} \mid \kw{false} $\\ \\
		\textit{Q. Commands} & $c$ & ::= &
		\begin{minipage}[t]{0.45\columnwidth}%
			$ \kw{mkQbit} M \mid \kw{measQbit} M \mid \kw{applyU} M$
		\end{minipage}\\
		\textit{Computations} & $E, F$ & ::= & $ \kw{return} M \mid x \leftarrow K; E \mid x \Leftarrow c; E \mid x =_A M; E $\\ \\
		\textit{Variable context} & $\Delta$ & ::= & $ \cdot \mid x : A, \Delta $ \\
		\textit{Q. Heap context} & $\Psi$ & ::= & $ \cdot \mid h, \Psi$ \\
		\textit{Assertion context} & $\Gamma$ & ::= & $ \cdot \mid P, \Gamma$ \\
	\end{tabular}
	\Description{Syntax of Quantum Hoare Type Theory}
	\caption{Syntax of Quantum Hoare Type Theory}
	\label{fig:syntax}
\end{figure}

@@

\paragraph{Quantum Commands}
These are the quantum-specific effectful commands from QIO for initialization, unitary application and measurement of qubits.

@@

\textbf{Code:}

@@

\textbf{Verification steps:}

\begin{minipage}{0.95\linewidth}
\begin{lstlisting}[language=QHaskell]
hqw : {emp} r : Bool {emp /\ Id(r, false)}
-- P0: emp
	= do q <= mkQbit false;
-- P1: P0 \o (q |-> |0\>)
		 measQbit q
-- P2: P1 \o ((q |-> -) -o emp)
\end{lstlisting}
\end{minipage}

@@

\paragraph{Quantum Language} Our programming language currently does not support conditionals or any iterative or recursive constructs. We would like to add those in future.
